Automatic Matching of Bullet Lands

### 2016-01-21 imaging-paper.tex
Include the submission folder

\documentclass[aoas, preprint]{imsart}

@@

\endlocaldefs

\begin{document}

@@

\linenumbers
%\tableofcontents
%\newpage
<<setup, fig.keep='all', cache=FALSE, echo=FALSE, eval=TRUE, message=F, warning=F>>=
#rm(list=ls())
#wd <- getwd()
library(extrafont)
library(knitr)
imgdir <- "Figure/"
codedir <- "code"
datadir <- "app/images/Hamby252_3DX3P1of2/"

options(replace.assign=TRUE,scipen=3, digits=2)
bstats <- read.csv("data-25-25/bullet-stats.csv")

scrubPath <- function(x) {
  splits <- strsplit(as.character(x), split="/")
  last <- sapply(splits, function(x) x[length(x)])
  gsub(".x3p","", last)
}

library(RColorBrewer)
library(ggplot2)
library(scales)
library(dplyr)
library(x3pr)
library(x3prplus)
library(grid)
library(gridExtra)
library(zoo)
library(tidyr)
library(rpart)
library(rpart.plot)
library(xtable)
library(sm)
@

@@

\begin{figure}[hbtp]
  \centering
<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2, fig.width=6, out.width='0.65\\textwidth'>>=

cols = c(alpha("grey60", alpha=0.6), alpha("black", 0.5))

br111 <- read.x3p(paste(datadir,"Br1 Bullet 1-5.x3p", sep = "/"))
dbr111 <- fortify_x3p(br111)

pars <- data.frame(getCircle(dbr111$y, dbr111$value))
dbr111$theta <- acos((dbr111$y-pars$x0)/pars$radius)/pi*180
dbr111 <- dbr111 %>% mutate(
  xpred = cos(theta/180*pi)*pars$radius + pars$x0,
  ypred = sin(theta/180*pi)*pars$radius + pars$y0
)

qplot(data=subset(dbr111, x <= 100*1.5625^2 & x >= 99*1.5625^2), y, value, geom="line", size=I(1)) +
  geom_line(aes(x=xpred, y=ypred, group=x), 
            colour="grey30", size=0.25) +
#  ylab(expression(paste("Surface Measurements (in ",mu,m,")", sep=""))) + 
  ylab("") +
  theme_bw() + 
  theme(legend.position="bottom") #+ coord_equal()
@
\caption{\label{fig:fixedX}Side profile of the surface measurements (in~$\mu m$) of a bullet land at a fixed height $x$. Note that the global features dominate any deviations, corresponding to the individual characteristics of striation marks.}
\end{figure}

@@

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[t]{\textwidth}\centering
\caption{\label{fig:loess_step1}Step 1 of identifying groove locations: For a fixed height ($x = 243.75\mu m$)  surface measurements for bullet~1-5 are plotted across the range of $y$.}
<<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE,  message=FALSE>>=
br111_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 1-5.x3p", sep = "/"))
br121_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-1.x3p", sep = "/"))
br122_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-2.x3p", sep = "/"))
br123_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-3.x3p", sep = "/"))
br124_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-4.x3p", sep = "/"))
br125_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-5.x3p", sep = "/"))
br126_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-6.x3p", sep = "/"))

p1 <- qplot(y, value, data = br111_bullet, size=I(1)) + theme_bw() + ylab("Surface measurements")
p1
@
\end{subfigure}
\begin{subfigure}[t]{\textwidth}\centering
\caption{\label{fig:loess_step3}Step 2 of identifying groove locations: The  surface measurements are smoothed twice with a smoothing factor of $s = 35$. The orange rectangle shows an example of the smoothing window. Valleys and peaks are detected, if they are not within the same window.}
<<loess_step3, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE,  message=FALSE>>=
subdata <- br111_bullet
smoothfactor <- 35

value_filled <- na.fill(subdata$value, "extend")
smoothed <- rollapply(value_filled, smoothfactor, function(x) mean(x))
smoothed_truefalse <- rollapply(smoothed, smoothfactor, function(x) mean(x))

xloc=250
p3 <- ggplot() +
  annotate("rect",xmin=xloc-17*1.5625, xmax=xloc+18*1.5625, ymin=80, ymax=200, fill=alpha("orange", alpha=0.5)) + 
  geom_point(aes(x=(1:length(smoothed_truefalse))*1.5625, y=smoothed_truefalse)) + theme_bw() +
    geom_vline(xintercept=xloc, colour="grey50") + xlab("y") + ylab("Surface measurements")

p3
@
\end{subfigure}
\begin{subfigure}[t]{\textwidth}\centering
\caption{\label{fig:loess_step47}Steps 3 -- 6 of identifying groove locations: After smoothing the surface measurements extrema on the left and right are detected (marked by vertical lines, red indicating peaks and blue indicating valleys). Values outside the blue boundaries are removed (shown in grey)}
<<loess_step47, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE,  message=FALSE>>=
    lengthdiff <- length(subdata$value) - length(smoothed_truefalse)

    peak_ind_smoothed <- head(which(rollapply(smoothed_truefalse, 3, function(x) which.max(x) == 2)), n = 1)
    peak_ind <- peak_ind_smoothed + floor(lengthdiff / 2)
    groove_ind <- head(which(rollapply(tail(smoothed_truefalse, n = -peak_ind_smoothed), 3, function(x) which.min(x) == 2)), n = 1) + peak_ind

    peak_ind2_smoothed_temp <- head(which(rollapply(rev(smoothed_truefalse), 3, function(x) which.max(x) == 2)), n = 1)
    peak_ind2_temp <- peak_ind2_smoothed_temp + floor(lengthdiff / 2)
    groove_ind2_temp <- head(which(rollapply(tail(rev(smoothed_truefalse), n = -peak_ind2_smoothed_temp), 3, function(x) which.min(x) == 2)), n = 1) + peak_ind2_temp

    peak_ind2 <- length(subdata$value) - peak_ind2_temp + 1
    groove_ind2 <- length(subdata$value) - groove_ind2_temp + 1

    p4 <- qplot(subdata$y[subdata$y < subdata$y[groove_ind]], subdata$value[subdata$y < subdata$y[groove_ind]], colour=I("grey60"), alpha=I(.25)) +
        ylim(c(min(subdata$value[!is.nan(subdata$value)]) - 25, max(subdata$value[!is.nan(subdata$value)]) + 25)) +
        xlim(c(min(subdata$y) - 25, max(subdata$y) + 25)) +
        geom_point(aes(x, y), data = data.frame(x = subdata$y[subdata$y > subdata$y[groove_ind2]], y = subdata$value[subdata$y > subdata$y[groove_ind2]]), colour=I("grey60"), alpha=I(.25)) +
        geom_point(inherit.aes = FALSE, aes(x, y), data = data.frame(x = subdata$y[subdata$y < subdata$y[groove_ind2] & subdata$y > subdata$y[groove_ind]], y = subdata$value[subdata$y < subdata$y[groove_ind2] & subdata$y > subdata$y[groove_ind]])) +
        theme_bw() +
        geom_vline(xintercept = subdata$y[peak_ind], colour = "red") +
        geom_vline(xintercept = subdata$y[groove_ind], colour = "blue") +
        geom_vline(xintercept = subdata$y[peak_ind2], colour = "red") +
        geom_vline(xintercept = subdata$y[groove_ind2], colour = "blue") +
      xlab("y") + ylab("Surface measurements")

    p4
@
\end{subfigure}
\caption{Overview of all six steps of the smoothing algorithm to identify and remove grooves from the bullet images.}
\end{figure}

@@

Figure~\ref{fig:loess_fit} shows the loess fit, in blue, overlaid on the processed image of bullet~1-5. The fit seems to do a reasonable job of capturing the structure of the image. The blue area around the signature are  95\% bootstrapped confidence intervals based on 1,000 bootstrap samples. 
Figure~\ref{fig:loess_resid} shows the residuals from this fit. These residuals are called the \emph{signature} of bullet~1-5.
%
\begin{figure}[hbtp]
  \centering
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{\label{fig:loess_fit} Loess fit for bullet~1-5.}
<<loess_fit, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\textwidth', warning=FALSE,  message=FALSE>>=

br111.groove <- get_grooves(br111_bullet)

my.loess <- fit_loess(br111_bullet, br111.groove)

my.loess$fitted + ylab("Surface measurements")
@
\end{subfigure}
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{\label{fig:loess_resid} Residuals of loess fit for bullet~1-5.
}
%<<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=
%my.loess$resid + ylab("Residuals of loess Fit") 
%@
<<boot_loess, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=
boot.loess <- boot_fit_loess(br111_bullet, br111.groove)
poly <- with(boot.loess, data.frame(x = c(y,rev(y)), y=c(high, rev(low))))
qplot(x=x, y=y, geom="polygon", fill=I("steelblue"), data=poly) +
  geom_line(aes(x=y, y=resid), data=my.loess$data) +
  ylab("Residuals of loess fit") + xlab("y") + theme_bw()
@
\end{subfigure}
\caption{Fit and residuals of a loess fit to bullet~1-5 (Barrel~1). The residuals define the {\it signature} of bullet~1-5. Blue areas around the residuals show  95\% bootstrap confidence intervals (based on $B=1,000$ bootstrap samples).
}
\end{figure}
%

@@

\subsection{Algorithm}\label{sec:algorithm}
<<twolands100, cache=TRUE,  echo=FALSE>>=
images <- file.path(datadir, dir(datadir))

#lof <- processBullets(paths = images[c(5,7)], x = 100)
subLOFx1 <- processBullets(read.x3p(images[5]), name = images[5], x = 100)
subLOFx2 <- processBullets(read.x3p(images[7]), name = images[7], x = 100)
#subLOFx1 <- subset(lof, bullet=="Br1 Bullet 1-5")
#subLOFx2 <- subset(lof, bullet=="Br1 Bullet 2-1")
subLOFx1$y <- subLOFx1$y + 23*1.5625 # working now!!!
lofX <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))

# smooth
lofX <- lofX %>% group_by(bullet) %>% mutate(
  l30 = smoothloess(y, resid, span = 0.03)
)

# cut at .75
threshold <- .75
lofX$r05 <- threshold* sign(lofX$l30) * as.numeric(abs(lofX$l30) > threshold)
lofX$type <- factor(lofX$r05)
levels(lofX$type) <- c("groove", NA, "peak")
@

<<twolands100adjust, echo=FALSE>>=
images <- file.path(datadir, dir(datadir))

#lof <- processBullets(paths = images[c(5,7)], x = 100)
subLOFx1 <- processBullets(read.x3p(images[5]), name = images[5], x = 100)
subLOFx2 <- processBullets(read.x3p(images[7]), name = images[7], x = 100)
#subLOFx1 <- subset(lof, bullet=="Br1 Bullet 1-5")
#subLOFx2 <- subset(lof, bullet=="Br1 Bullet 2-1")
subLOFx1$y <- subLOFx1$y - min(subLOFx1$y)
subLOFx2$y <- subLOFx2$y - min(subLOFx2$y)
ccf <- ccf(subLOFx1$resid, subLOFx2$resid, lag.max=100, plot=FALSE)
lag <- ccf$lag[which.max(ccf$acf)]
subLOFx1$y <- subLOFx1$y - lag*1.5625 
lofY <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))

@

<<twolands100match, cache=TRUE, echo=FALSE, dependson='twolands100'>>=
matches <- lofX %>% group_by(y) %>% summarise(
  potential = (length(unique(type)) == 1),
  allnas = sum(is.na(type))/n(),
  type1 = na.omit(type)[1],
  type = paste(type, sep="|", collapse="|"),
  n = n()
)

matches$id <- cumsum(matches$allnas == 1) + 1
matches$lineid <- as.numeric(matches$allnas != 1) * matches$id

isMatch <- function(id, type) {
  if (id[1] == 0) return(FALSE)
#  browser()
  types <- strsplit(type, split = "|", fixed=TRUE) 
  t1 <- sapply(types, function(x) x[1])
  t2 <- sapply(types, function(x) x[2])
  if (all(t1 == "NA")) return(FALSE)
  if (all(is.na(t2))) return(FALSE)
  if (all(t2 == "NA")) return(FALSE)

  peak <- length(grep("peak", c(t1, t2))) > 0
  groove <- length(grep("groove", c(t1, t2))) > 0
  if (peak & groove) return(FALSE)

  return(TRUE)
}

lines <- matches %>% group_by(lineid) %>% summarise(
  meany = mean(y, na.rm=T),
  miny = min(y, na.rm=T),
  maxy = max(y, na.rm=T) + 1.5625,
  match = isMatch(lineid, type),
  type = type1[1]
)
lines <- subset(lines, lineid != 0)
@

\begin{figure}[hbtp]
\centering
    \begin{subfigure}[t]{\textwidth}\centering
\caption{Loess smooth of signatures  at a height of $x = 100\mu m$ (span is 0.03).\label{fig:smooth}}{%
<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.65\\textwidth', warning=FALSE>>=
lofX$bulletname <- scrubPath(lofX$bullet)
qplot(data=lofX, y, resid, colour=I("grey30"), size=I(0.75), geom="line", group=bulletname) +
  geom_line(aes(y=l30), colour="grey70", size=0.5) +
  facet_grid(bulletname~.) +
#  scale_colour_manual("", values=cols) +
  theme_bw() + ylab("") + 
  theme(legend.position="none")
@
    }
\end{subfigure}
\begin{subfigure}[t]{\textwidth}\centering
\caption{Using a rolling median peaks and valleys are identified for each signature. Peaks and valleys on the signature correspond to striation marks on the bullet's surface. \label{fig:smoothcutb}}{%
<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.65\\textwidth', warning=FALSE>>=
#images <- file.path(datadir, dir(datadir))
  #lof <- processBullets(paths = images[c(5,7)], x = 100)

  subLOFx1 <- processBullets(read.x3p(images[5]), name = images[5], x = 100)
  subLOFx2 <- processBullets(read.x3p(images[7]), name = images[7], x = 100)
  lof <- rbind(subLOFx1, subLOFx2)

  lof <- bulletSmooth(lof)
  bAlign = bulletAlign(lof)
  lofX <- bAlign$bullet

  b12 <- unique(lof$bullet)
  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = 25)
  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 25)
  peaks1$lines$bullet <- b12[1]
  peaks2$lines$bullet <- b12[2]
  peaks <- rbind(peaks1$lines, peaks2$lines)

  peaks$bulletname <- scrubPath(peaks$bullet)
  lofX$bulletname <- scrubPath(lofX$bullet)
  ggplot() + theme_bw() +
    geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, 
              data=peaks,  alpha=0.2) +
    geom_vline(aes(xintercept=extrema, colour=factor(type)), 
               data= peaks, alpha=0.7) +
    scale_colour_brewer(palette="Set2") + 
    scale_fill_brewer(palette="Set2") +
    theme(legend.position="none") + 
    facet_grid(bulletname~.) +
    geom_line(aes(x=y, y=l30, group=bulletname), data=lofX) +
  ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) 
@
}
\end{subfigure}    
\begin{subfigure}[t]{\textwidth}\centering
    \caption{Rectangles in the back identify a striation mark on one of the bullets.  Matching striation marks are indicated by color filled rectangles and marked by an `o'. Mismatches are filled in grey and  marked by an `x'.   \label{fig:smoothcutd}}{%
<<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='.65\\textwidth', warning=FALSE>>=
peaks1$lines$bullet <- b12[1]
peaks2$lines$bullet <- b12[2]

lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)

ggplot() + 
  geom_rect(aes(xmin = xmin, xmax = xmax, fill=factor(type)), ymin = -6, ymax=6.5,  data = lines, alpha=0.2, show.legend = FALSE) +
  theme(legend.position="bottom") +
  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(lines, !match)) +
  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(lines, match)) +
  ylim(c(-6,6.5)) + theme_bw() +
  geom_line(data=lofX, aes(x=y, y=l30, group=bulletname, linetype=bulletname)) +
  scale_linetype_discrete("") +
  scale_colour_manual("", values=cols) +
  scale_fill_brewer("", palette="Set2", na.value=alpha("grey60", 0.5)) +
    theme(legend.position = c(1,1.1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) +
  xlab("y")
@
    }
\end{subfigure}
\caption{\label{fig:match}Matching striation marks: smooth (a), identify peaks and valley (b), and match peaks and valleys between signatures (c).}
\end{figure}
%\afterpage{\clearpage}
Figure~\ref{fig:match} gives an overview of the automated matching routine: 
We first identify a stable region for each bullet land and extract the signature at the lowest height in this region, because typically, individual characteristics are best expressed at the lower end of the bullet (see Supplement section~\ref{supp:bulletbottom} for a more detailed discussion).

@@

\subsection{Horizontal alignment}
Signatures of each of the two lands, 1-5 and 2-1,  in Figure~\ref{fig:manualmatch-rgl} are shown in Figure~\ref{fig:cross100} extracted at a height of $x = 100\mu m$. Striation marks show up in this representations as peaks and valleys.  The individual characteristics are quite prominent and, again, suggest a match between the lands. A horizontal shift of one of the signatures (result shown in Fig~\ref{fig:crosscutX}) emphasizes the strong similarities. 
%
\begin{figure}[hbtp]
\centering
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{Raw bullet land signatures.\label{fig:crosscut}}{%
<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
lof$bulletname <- scrubPath(lof$bullet)
qplot(y, resid, group=bulletname, colour=bulletname, data=lof, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(1, 1.1), legend.justification=c("right", "top"),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}    
\begin{subfigure}[b]{.49\textwidth}\centering
    \caption{Aligned signatures.\label{fig:crosscutX}}{%
<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
lofY$bulletname <- scrubPath(lofY$bullet)
qplot(y+min(lof$y), resid, group=bulletname, colour=bulletname, data=lofY, 
      geom="line") + xlab("y") +
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(1, 1.1), legend.justification=c("right", "top"),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}
\caption{\label{fig:cross100}Signatures of bullets 1-5 and 2-1 taken at  heights of $x = 100\mu m$. A horizontal shift of the values of bullet~1-5 to the right shows the similarity of the striation marks.}
\end{figure}
%
This horizontal shift is based on the cross-correlation between the two signatures: let $f(t)$ and $g(t)$ define the signature values  at $t$, where $t$ are locations between 0~$\mu m$ and about 2500~$\mu m$, 1.5625~$\mu m$ apart.
The cross-correlation between $f$ and $g$ at lag $k$ is then defined as
\[
(f * g) (k) = \sum_t f(t+k) g(t),
\]
with suitably defined limits for the summation. % (because variances of $f$ and $g$ do not change, the above equation is proportional to the correlation).

\begin{figure}[hbtp]
  \centering
<<ccf, echo=FALSE, fig.width = 8, fig.height = 3, out.width = '.65\\textwidth'>>=
lag <- ccf$lag[which.max(ccf$acf)]
qplot(x=ccf$lag, xend=ccf$lag, y = 0, yend= ccf$acf, geom="segment", colour = I("grey50")) + 
  theme_bw() + ylab("Correlation") + xlab("Lag k") +
  geom_segment(aes(x = lag, xend = lag, y = 0, yend =  max(ccf$acf)), colour="black") + 
  scale_x_continuous(breaks = c(-100, -50, lag, 0, 50, 100),
                     minor_breaks = c(-75,-25,25, 75))
@
\caption{\label{fig:ccf}Cross-correlation function between the two signatures shown in Figure~\ref{fig:crosscut} at lags between -100 and 100. At a lag of -17 the correlation peaks, indicating the largest amount of agreement between the signatures. Figure~\ref{fig:crosscutX} shows the lag-shifted signatures.}
\end{figure}

@@

\begin{figure}[hbtp]
  \centering
<<crosscuts-vary-b31, echo=FALSE, fig.width = 12, fig.height = 7, out.width = '\\textwidth', cache=TRUE, warning = FALSE>>=
paths <- file.path(datadir, dir(datadir))
crosscuts <- seq(100,250, by=25)
lof <- processBullets(read.x3p(paths[37]), name = paths[37], x = crosscuts)
lof$bullet <- paste(scrubPath(lof$bullet), lof$x)

reslist <- lapply(1:length(crosscuts[-1]), function(i) {
  b2 <- subset(lof, x %in% crosscuts[i:(i+1)])
  lofX <- bulletSmooth(b2)
  bAlign <- bulletAlign(lofX)
  lofX <- bAlign$bullet

  b12 <- unique(b2$bullet)
  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = 25)
  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 25)

>>>>

\documentclass[aoas, preprint]{imsart}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}

@@

\endlocaldefs
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

@@

\linenumbers
%\tableofcontents
%\newpage

@@

\begin{figure}[hbtp]
  \centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=0.65\textwidth]{figure/fixedX-1}

\end{knitrout}
\caption{\label{fig:fixedX}Side profile of the surface measurements (in~$\mu m$) of a bullet land at a fixed height $x$. Note that the global features dominate any deviations, corresponding to the individual characteristics of striation marks.}
\end{figure}

@@

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[t]{\textwidth}\centering
\caption{\label{fig:loess_step1}Step 1 of identifying groove locations: For a fixed height ($x = 243.75\mu m$)  surface measurements for bullet~1-5 are plotted across the range of $y$.}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=.5\textwidth]{figure/loess_step1-1}

\end{knitrout}
\end{subfigure}
\begin{subfigure}[t]{\textwidth}\centering
\caption{\label{fig:loess_step3}Step 2 of identifying groove locations: The  surface measurements are smoothed twice with a smoothing factor of $s = 35$. The orange rectangle shows an example of the smoothing window. Valleys and peaks are detected, if they are not within the same window.}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=.5\textwidth]{figure/loess_step3-1}

\end{knitrout}
\end{subfigure}
\begin{subfigure}[t]{\textwidth}\centering
\caption{\label{fig:loess_step47}Steps 3 -- 6 of identifying groove locations: After smoothing the surface measurements extrema on the left and right are detected (marked by vertical lines, red indicating peaks and blue indicating valleys). Values outside the blue boundaries are removed (shown in grey)}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=.5\textwidth]{figure/loess_step47-1}

\end{knitrout}
\end{subfigure}
\caption{Overview of all six steps of the smoothing algorithm to identify and remove grooves from the bullet images.}
\end{figure}

@@

Figure~\ref{fig:loess_fit} shows the loess fit, in blue, overlaid on the processed image of bullet~1-5. The fit seems to do a reasonable job of capturing the structure of the image. The blue area around the signature are  95\% bootstrapped confidence intervals based on 1,000 bootstrap samples. 
Figure~\ref{fig:loess_resid} shows the residuals from this fit. These residuals are called the \emph{signature} of bullet~1-5.
%
\begin{figure}[hbtp]
  \centering
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{\label{fig:loess_fit} Loess fit for bullet~1-5.}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\textwidth]{figure/loess_fit-1}

\end{knitrout}
\end{subfigure}
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{\label{fig:loess_resid} Residuals of loess fit for bullet~1-5.
}
%<<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=
%my.loess$resid + ylab("Residuals of loess Fit") 
%@
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\textwidth]{figure/boot_loess-1}

\end{knitrout}
\end{subfigure}
\caption{Fit and residuals of a loess fit to bullet~1-5 (Barrel~1). The residuals define the {\it signature} of bullet~1-5. Blue areas around the residuals show  95\% bootstrap confidence intervals (based on $B=1,000$ bootstrap samples).
}
\end{figure}
%

@@

\subsection{Algorithm}\label{sec:algorithm}

\begin{figure}[hbtp]
\centering
    \begin{subfigure}[t]{\textwidth}\centering
\caption{Loess smooth of signatures  at a height of $x = 100\mu m$ (span is 0.03).\label{fig:smooth}}{%
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=.65\textwidth]{figure/smooth-1}

\end{knitrout}
    }
\end{subfigure}
\begin{subfigure}[t]{\textwidth}\centering
\caption{Using a rolling median peaks and valleys are identified for each signature. Peaks and valleys on the signature correspond to striation marks on the bullet's surface. \label{fig:smoothcutb}}{%
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=.65\textwidth]{figure/smoothcut-1}

\end{knitrout}
}
\end{subfigure}    
\begin{subfigure}[t]{\textwidth}\centering
    \caption{Rectangles in the back identify a striation mark on one of the bullets.  Matching striation marks are indicated by color filled rectangles and marked by an `o'. Mismatches are filled in grey and  marked by an `x'.   \label{fig:smoothcutd}}{%
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=.65\textwidth]{figure/smoothmatch-1}

\end{knitrout}
    }
\end{subfigure}
\caption{\label{fig:match}Matching striation marks: smooth (a), identify peaks and valley (b), and match peaks and valleys between signatures (c).}
\end{figure}
%\afterpage{\clearpage}
Figure~\ref{fig:match} gives an overview of the automated matching routine: 
We first identify a stable region for each bullet land and extract the signature at the lowest height in this region, because typically, individual characteristics are best expressed at the lower end of the bullet (see Supplement section~\ref{supp:bulletbottom} for a more detailed discussion).

@@

\subsection{Horizontal alignment}
Signatures of each of the two lands, 1-5 and 2-1,  in Figure~\ref{fig:manualmatch-rgl} are shown in Figure~\ref{fig:cross100} extracted at a height of $x = 100\mu m$. Striation marks show up in this representations as peaks and valleys.  The individual characteristics are quite prominent and, again, suggest a match between the lands. A horizontal shift of one of the signatures (result shown in Fig~\ref{fig:crosscutX}) emphasizes the strong similarities. 
%
\begin{figure}[hbtp]
\centering
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{Raw bullet land signatures.\label{fig:crosscut}}{%
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\textwidth]{figure/crosscuts-1}

\end{knitrout}
    }
\end{subfigure}    
\begin{subfigure}[b]{.49\textwidth}\centering
    \caption{Aligned signatures.\label{fig:crosscutX}}{%
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\textwidth]{figure/crosscutsX-1}

\end{knitrout}
    }
\end{subfigure}
\caption{\label{fig:cross100}Signatures of bullets 1-5 and 2-1 taken at  heights of $x = 100\mu m$. A horizontal shift of the values of bullet~1-5 to the right shows the similarity of the striation marks.}
\end{figure}
%
This horizontal shift is based on the cross-correlation between the two signatures: let $f(t)$ and $g(t)$ define the signature values  at $t$, where $t$ are locations between 0~$\mu m$ and about 2500~$\mu m$, 1.5625~$\mu m$ apart.
The cross-correlation between $f$ and $g$ at lag $k$ is then defined as
\[
(f * g) (k) = \sum_t f(t+k) g(t),
\]
with suitably defined limits for the summation. % (because variances of $f$ and $g$ do not change, the above equation is proportional to the correlation).

\begin{figure}[hbtp]
  \centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=.65\textwidth]{figure/ccf-1}

\end{knitrout}
\caption{\label{fig:ccf}Cross-correlation function between the two signatures shown in Figure~\ref{fig:crosscut} at lags between -100 and 100. At a lag of -17 the correlation peaks, indicating the largest amount of agreement between the signatures. Figure~\ref{fig:crosscutX} shows the lag-shifted signatures.}
\end{figure}

@@

\begin{figure}[hbtp]
  \centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\textwidth]{figure/crosscuts-vary-b31-1}

\end{knitrout}
\caption{\label{fig:crosscuts2}Signatures  for barrel 3, bullet~1-1 extracted from varying heights. Initially, the match between signatures taken at heights 25$\mu m$ apart is affected strongly by some break off at the bottom of the bullet. At a  level of $175\mu m$ the bullet's signature stabilizes. For this land, matches should not be attempted at lower heights. }
\end{figure}
%An opportunity that arises from comparing multiple cross sections of a bullet's surface measurements is the following:
by comparing signatures from heights that are not too far apart --  25~$\mu m$ to 50~$\mu m$ -- we get an indication whether the signatures come from a rapidly changing section of the surface, indicative of a break-off or some other damage, or from a stable section, where we have a reasonable expectation of finding matches to other signatures. In the approach here, we keep increasing the height $x$ at which the signature is taken until we find a section with a stable pattern. This process is shown in Figure~\ref{fig:crosscuts2} at the example of bullet~1-1 from barrel 3, where
`stability' is  defined as two aligned signatures from heights chosen 25$\mu m$ apart having a cross-correlation of at least 0.95.
